
\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{fancyhdr}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}

% GTU Colors
\definecolor{gtublue}{RGB}{0,51,102}
\definecolor{gtuorange}{RGB}{255,102,0}

% Header and Footer
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\textcolor{gtublue}{\textbf{Gujarat Technological University}}}
\fancyhead[R]{\textcolor{gtuorange}{\textbf{COGC-2021}}}
\fancyfoot[C]{\thepage}

% Section styling
\titleformat{\section}{\Large\bfseries\color{gtublue}}{}{0em}{}
\titleformat{\subsection}{\large\bfseries\color{gtuorange}}{}{0em}{}

% Custom commands
\newcommand{\courseheader}[2]{
    \begin{center}
        \textcolor{gtublue}{\Huge\textbf{#1}}\\
        \vspace{0.3cm}
        \textcolor{gtuorange}{\Large Course Code: #2}\\
        \vspace{0.5cm}
        \rule{\textwidth}{2pt}
    \end{center}
}

\begin{document}
\courseheader{Python Programming}{DI01016011}


\section{Course Information}

\begin{tabular}{|l|l|}
\hline
\textbf{Field} & \textbf{Details} \\
\hline
Program & Diploma in Engineering \\
\hline
Branch & Information Technology \\
\hline
Level & Diploma \\
\hline
Semester & 1 \\
\hline
Academic Year & 2024-2025 \\
\hline
Category & PCC \\
\hline
Prerequisites & Basic computer skills, including the ability to write basic statements and expressions. \\
\hline
\end{tabular}


\section{Rationale}

Computer programming skills are now becoming part of basic education as these skills are increasing of vital importance for future job and career prospects. The Python programming language is one of the most popular programming languages worldwide. The course emphasizes the use of python programming in multiple domains. Python is a modern language for writing compact codes specifically for programming Server-side web apps, Data Analytics and Machine Learning, an important Artificial Intelligence domain. Furthermore, Python has gained popularity in scientific computing, production tools and game programming. This course focuses on developing python programming to do a variety of programming tasks where the students are encouraged to develop basic applications using different open source tools. At the end of the course, the student will be developing adequate basic programming skills using python language.


\section{Course Outcomes}

After completion of the course, students will be able to:

\begin{longtable}{|p{1cm}|p{12cm}|p{2cm}|}
\hline
\textbf{No.} & \textbf{Course Outcomes} & \textbf{RBT Level} \\
\hline
\endhead
CO1 & Prepare flowchart and algorithm for solving computing problems. & Apply \\
\hline
CO2 & Develop python programs to solve simple problems. & Apply \\
\hline
CO3 & Apply control structure feature of python for developing programs. & Apply \\
\hline
CO4 & Develop programs in Python using built-in functions, modules, and library functions. & Apply \\
\hline
CO5 & Develop python programs applying strings and lists manipulation concepts. & Apply \\
\hline
\end{longtable}


\section{Teaching and Examination Scheme}

\begin{center}
\begin{tabular}{|c|c|c|c||c|c|c|c|c|}
\hline
\multicolumn{4}{|c|}{\textbf{Teaching Scheme (Hours)}} & \multicolumn{5}{c|}{\textbf{Assessment Pattern (Marks)}} \\
\hline
\textbf{L} & \textbf{T} & \textbf{PR} & \textbf{C} & \textbf{Theory ESE} & \textbf{Theory CA} & \textbf{Practical CA} & \textbf{Practical ESE} & \textbf{Total} \\
\hline
3 & 0 & 2 & 4 & 70 & 30 & 20 & 30 & 150 \\
\hline
\end{tabular}
\end{center}


\section{Course Content}

\begin{longtable}{|p{2cm}|p{10cm}|p{2cm}|p{2cm}|}
\hline
\textbf{Unit No.} & \textbf{Content} & \textbf{Hours} & \textbf{Weightage (\%)} \\
\hline
\endhead
1 & \textbf{Problem Solving using Flowchart and Algorithm}\\
1.1 Introduction, Steps for problem-solving, Algorithm and its characteristics, Importance of algorithm.\\
1.2 Symbolic representation of a flowchart, Importance and Limitations of flowchart, Flow of control\\
1.3 Problem solving using pseudocode\\
 & 5 & 11 \\
\hline
2 & \textbf{Basics of Python}\\
2.1 Introduction to python, Python features, Applications of python programming\\
2.2 Python installation\\
2.3 Basic structure of python program, Python Comments, Keywords, identifiers, variables, Data types, and Operators.\\
2.4 Type Conversion\\
 & 10 & 17 \\
\hline
3 & \textbf{Flow of Control}\\
3.1 Introduction to Flow of Control\\
3.2 Selection\\
• If statement\\
• Elif statement\\
• Nested if statement\\
• match statement\\
3.3 Repetition\\
• For loop\\
• While loop\\
• Nested loop\\
3.4 break, continue, and pass Statements\\
 & 10 & 24 \\
\hline
4 & \textbf{Functions}\\
4.1 Introduction to Functions\\
4.2 User Defined Functions\\
• Arguments and Parameters\\
4.3 Scope of a Variable\\
• Global Variable\\
• Local Variable\\
Python Standard Library\\
• Built-in functions\\
• Input or output - input(), print()\\
• Mathematical Functions - abs(), divmod(), max(), min(), pow(), sum()\\
• Module\\
  ○ math\\
  ○ random\\
  ○ statistics\\
 & 10 & 24 \\
\hline
5 & \textbf{Strings and Lists}\\
5.1 Introduction to Strings, String Operations, Traversing a String\\
5.2 Strings Methods and Built-in Functions\\
5.3 Introduction to List and its Operations\\
5.4 List Methods and Built-in Functions\\
• Nested and Copying Lists\\
5.5 List as Arguments to Function\\
 & 10 & 24 \\
\hline
\end{longtable}


\section{Suggested Course Practical List}

\begin{longtable}{|p{1cm}|p{10cm}|p{2cm}|p{2cm}|}
\hline
\textbf{Sr. No} & \textbf{Practical Outcomes (PrOs)} & \textbf{Unit No.} & \textbf{Hrs.} \\
\hline
\endhead
1 & Prepare flowchart and algorithm for a given problem.\\ • Find the sum of two given numbers.\\ • Find a maximum out of two given numbers.\\ • Find whether a given number is odd or even.\\ • Find a maximum out of three given numbers. & 1 & 2 \\
\hline
2 & Install \textbackslash{}& configure python software and Create a program to print your name, date of birth and mobile number. & 2 & 2 \\
\hline
3 & Develop a program to identify data-types in python. & 2 & 2 \\
\hline
4 & Create programs for mathematical operations and conversions.\\ • 1) Create a program to read three numbers from the user and find the average of the numbers.\\ • 2) Create a program to convert temperature from Fahrenheit to Celsius unit using eq: C=(F-32)/1.8 & 2 & 2 \\
\hline
5 & Create programs for conditional statements and comparison operations.\\ • 1) Create a program to identify whether the scanned number is even or odd and print an appropriate message.\\ • 2) Create a program to find a maximum number among the given three numbers. & 3 & 2 \\
\hline
6 & Develop a program to show whether the entered number is prime or not. & 3 & 2 \\
\hline
7 & Develop a program to print odd and even numbers from 1 to N numbers. (Where N is an integer number entered by the user) & 3 & 2 \\
\hline
8 & Develop a program to demonstrate the use of break, continue and pass statements. & 3 & 2 \\
\hline
9 & Develop user-defined functions for mathematical operations.\\ • 1) Develop a user-defined function to find the factorial of a given number.\\ • 2) Create a user-defined function to print the Fibonacci series of 0 to N numbers. (Where N is an integer number and passed as an argument) & 4 & 2 \\
\hline
10 & Write a program using the function that reverses the entered value. & 4 & 2 \\
\hline
11 & Write a program that determines whether a given number is an Armstrong number or not using a user-defined function. & 4 & 2 \\
\hline
12 & Write programs for string manipulation operations.\\ • 1) Write a program to reverse words in a given sentence.\\ • 2) Write a program to check if a substring is present in a given string.\\ • 3) Write a program to count and display the number of vowels, consonants, uppercase, lowercase characters in a string. & 5 & 2 \\
\hline
13 & Create programs for list operations and analysis.\\ • 1) Create a program to find the sum of all elements in a list using a loop.\\ • 2) Create a program to find the smallest and largest element in a given list. & 5 & 3 \\
\hline
14 & Given a list saved in variable: a = [1, 8, 7, 15, 25, 36, 48, 64, 81, 95]. Write a Python program that takes this list and makes a new list that has only the even elements of this list in it. & 5 & 3 \\
\hline
\end{longtable}


\section{Suggested Specification Table with Marks (Theory)}

\begin{center}
\begin{tabular}{|l|l|c|c|c|c|c|c|c|}
\hline
\textbf{Unit} & \textbf{Unit Title} & \textbf{R} & \textbf{U} & \textbf{A} & \textbf{N} & \textbf{E} & \textbf{C} & \textbf{Total} \\
\hline
Overall Distribution & Theory Marks Distribution & 23 & 28 & 49 & 0 & 0 & 0 & 100 \\
\hline
\end{tabular}
\end{center}

\textbf{Legend:} R: Remember; U: Understanding; A: Application; N: Analyze; E: Evaluate; C: Create


\section{References/Suggested Learning Resources}


\subsection{Books}

\begin{longtable}{|p{1cm}|p{6cm}|p{4cm}|p{4cm}|}
\hline
\textbf{Sr. No.} & \textbf{Title of Book} & \textbf{Author} & \textbf{Publication} \\
\hline
\endhead
1 & Learn Programming in Python with Cody Jackson & Cody Jackson & Packt Publishing, 2018 \\
\hline
2 & Python Basics: A Practical Introduction to Python 3 & David Amos, Dan Bader et. al. & Real Python, 2021 \\
\hline
3 & Introduction to Problem Solving with Python & E. Balagurusamy & Mc Graw Hill India, New Delhi \\
\hline
4 & Beginning Python & James Payne & Wiley, 2010 \\
\hline
5 & Think Python & Allen Downey & O'Reilly, USA, 2016 \\
\hline
\end{longtable}

\subsection{Open-source Software and Websites}

\begin{enumerate}
\item Python.org - \url{https://www.python.org} - Official Python website
\item Learn Python.org - \url{https://www.learnpython.org} - Interactive Python tutorial
\item Hackr.io Python Tutorials - \url{https://www.hackr.io/tutorials/learn-python} - Curated Python learning resources
\item SoloLearn Python Course - \url{https://www.sololearn.com/learning/1073} - Mobile-friendly Python course
\item NPTEL - \url{https://www.nptel.iitm.ac.in} - National Programme on Technology Enhanced Learning
\end{enumerate}


\section{Suggested Project List}

\begin{enumerate}
\item \textbf{Number Systems Converter} - Develop a console based application to convert number systems (i.e., Binary to Decimal, Binary to Octal, etc.)
\item \textbf{Simple Calculator} - Develop a simple calculator.
\item \textbf{Countdown Timer} - Develop console based Countdown timer.
\item \textbf{Number System Table Generator} - Print number system table for given range (Decimal | Binary | Octal | Hexadecimal)
\item \textbf{Fizz-Buzz Game} - Develop a fizz-buzz game.
\item \textbf{Password Generator} - Develop a program to generate random passwords (8 Digits).
\end{enumerate}


\section{Suggested Activities for Students}

\begin{enumerate}
\item Prepare a document which differentiates python versions. (\textit{Research})\\
Students should perform following activities in group and prepare reports of about 5 pages for each activity. They should also collect/record physical evidences for their (student's) portfolio which may be useful for their placement interviews.
\item Undertake projects in teams (\textit{Collaborative Project})\\
Team-based project work to enhance collaboration skills and practical application of learned concepts.
\item Give a seminar on any relevant topics. (\textit{Presentation})\\
Individual or group presentations to develop communication skills and deepen understanding of Python-related topics.
\item Undertake a market survey of different python frameworks. (\textit{Market Research})\\
Industry analysis to understand current trends and applications of Python frameworks in the market.
\end{enumerate}


\section{Laboratory/Learning Resources Required}

\begin{longtable}{|p{1cm}|p{10cm}|p{3cm}|}
\hline
\textbf{Sr. No.} & \textbf{Laboratory/Learning Resources} & \textbf{Applicable To} \\
\hline
\endhead
1 & \textbf{Computer System}\\
Operating system: Windows 7 or higher Ver., macOS, and Linux, with 4GB or higher RAM, Python versions: 2.7.X, 3.6.X, or higher Ver. & All PrOs \\
\hline
2 & \textbf{Python IDEs and Code Editors}\\
Open Source: IDLE, Jupyter & PrO 2 to 14 \\
\hline
\end{longtable}

\end{document}